---
title: "ST2195 Question 1 to 5 - ID: 200615961"
output: html_document
---
Installing the packages
```{r}
#install.packages("RSQLite")
#install.packages("DBI")
#install.packages("dplyr")
#install.packages("data.table")
#install.packages("lubridate")
#install.packages("tibble")
#install.packages("stringi")
#install.packages("ggplot2")
#install.packages("scales")
#install.packages("ggpubr")
#install.packages("mlr3")
#install.packages("mlr3learners")
#install.packages("mlr3pipelines")
#install.packages("mlr3tuning")
#install.packages("mlr3viz")
#install.packages("knitr")
#install.packages("broom")
#install.packages("xgboost")
#install.packages("e1071")
#install.packages("caret")
#install.packages("InformationValue")
#install.packages("ISLR")
#install.packages("pROC")
```

Set the libraries
```{r}
library(knitr)
library(RSQLite)
library(DBI)
library(dplyr)
library(data.table)
library(lubridate)
library(tibble)
library(stringi)
library(ggplot2)
library(scales)
library(ggpubr)
library(ggrepel)
library(mlr3)
library(mlr3learners)
library(mlr3pipelines)
library(mlr3tuning)
library(mlr3viz)
library(e1071)
library(broom)
library(caret)
library(InformationValue)
library(ISLR)
library(pROC)
```

Set the working directory and knitr for easier reading
```{r}
setwd("/Users/malcolmteo/Desktop/Programming Exam")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Set the base
````{r}
#Create a new SQL database called airline2_r 
base <- dbConnect(RSQLite::SQLite(), "airline2_r.db")

#Read the various csv files in the folder
airports <- read.csv("airports.csv", header = TRUE)
carriers <- read.csv("carriers.csv", header = TRUE)
planes <- read.csv("plane-data.csv", header = TRUE)

#Upload the previous three tables into the databaase
dbWriteTable(base, "airports", airports, overwrite = TRUE)
dbWriteTable(base, "carriers", carriers, overwrite = TRUE)
dbWriteTable(base, "planes", planes, overwrite = TRUE)

#Read the various csv files in the folder
a2004 <- read.csv("2004.csv", header = TRUE)
a2005 <- read.csv("2005.csv", header = TRUE)

#Combine the two new data frames - a2004 and a2005 - into a single data frame called alpha
alpha <- union(a2004,a2005)

#Remove na values in alpha and planes data frame
alpha <- na.omit(alpha)
planes <- na.omit(planes)

#Save ram space
rm(a2004) 
rm(a2005)
```

Specify Day, Month, Time of Day, Time of Year, Punctuality 
```{r}
#Use conditions to create new columns within the data frame
alpha$DepTimeofDay <-  ifelse(alpha$DepTime <= 600, "Midnight",
                      ifelse(alpha$DepTime <= 1200, "Morning",
                      ifelse(alpha$DepTime <= 1800, "Afternoon",
                      ifelse(alpha$DepTime <= 2400, "Night",
                      ifelse(alpha$DepTime <= 3000, "Midnight",
"Wrong")))))

alpha$Day <-  ifelse(alpha$DayOfWeek == 1, "Mon",
              ifelse(alpha$DayOfWeek == 2, "Tues",
              ifelse(alpha$DayOfWeek == 3, "Wed",
              ifelse(alpha$DayOfWeek == 4, "Thurs",
              ifelse(alpha$DayOfWeek == 5, "Fri",
              ifelse(alpha$DayOfWeek == 6, "Sat",
              ifelse(alpha$DayOfWeek == 7, "Sun",
"Wrong")))))))

alpha$TimeofYear <- ifelse(alpha$Month <=3, "First Quarter",
                    ifelse(alpha$Month <=6, "Second Quarter",
                    ifelse(alpha$Month <=9, "Third Quarter",
                    ifelse(alpha$Month <=12, "Fourth Quarter",
"Wrong"))))

alpha$Mth <-  ifelse(alpha$Month == 1, "Jan",
                    ifelse(alpha$Month == 2, "Feb",
                    ifelse(alpha$Month == 3, "Mar",
                    ifelse(alpha$Month == 4, "Apr",
                    ifelse(alpha$Month == 5, "May",
                    ifelse(alpha$Month == 6, "Jun",
                    ifelse(alpha$Month == 7, "Jul",
                    ifelse(alpha$Month == 8, "Aug",
                    ifelse(alpha$Month == 9, "Sep",
                    ifelse(alpha$Month == 10, "Oct",
                    ifelse(alpha$Month == 11, "Nov",
                    ifelse(alpha$Month == 12, "Dec",
"Wrong"))))))))))))

alpha$Punctuality <- ifelse(alpha$ArrDelay <=0, "Early",
                            "Late")

# Append into DBI
dbWriteTable(base, "alpha", alpha, overwrite = TRUE)
```

Question 1:
Best time of the day & year to fly
```{r}
#Create a bar plot for best time of day to fly referencing the departure delay
alpha %>%
  mutate(DepTimeofDay = factor(x=DepTimeofDay,
                               levels=c("Morning", "Afternoon", 
                                        "Night", "Midnight"))) %>%
  ggplot(aes(x=DepTimeofDay, 
             y= DepDelay)) +
  stat_summary (fun = "mean",
                geom = "bar",
                fill = "blue", 
                alpha = 0.7) +
  theme_classic() +
  ggtitle("Time of Day against Average Departure Delay") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time of Day") +
  ylab("Average Departure Delay")

#The fourth quarter is least delayed on average at 34.6 mins
#This is considering when there is departure and arrival Delay
Best_TimeGivenADDD <- dbGetQuery(base,
  "SELECT TimeofYear as TimeofYear, AVG(DepDelay) as Avg_DepDelay
  FROM alpha
  WHERE alpha.DepDelay >0 AND alpha.ArrDelay >0
  GROUP BY TimeofYear
  ORDER by Avg_DepDelay"
)

#The fourth quarter is least delayed on average at 28.54 mins
#This is considering when there is only departure delay
Best_TimeGivenDD <- dbGetQuery(base,
  "SELECT TimeofYear as TimeofYear, AVG(DepDelay) as Avg_DepDelay
  FROM alpha
  WHERE alpha.DepDelay >0
  GROUP BY TimeofYear
  ORDER by Avg_DepDelay"
)

#The second quarter is least delayed on average at 7.59 mins
#This is when considering the whole population
Best_Time <- dbGetQuery(base,
  "SELECT TimeofYear as TimeofYear, AVG(DepDelay) as Avg_DepDelay
  FROM alpha
  GROUP BY TimeofYear
  ORDER by Avg_DepDelay"
)

#The second quarter, months april, may, june, has the lowest departure delay
Best_Time %>%
  mutate(TimeofYear = factor(TimeofYear, 
                            levels = c("First Quarter", "Second Quarter",
                                       "Third Quarter", "Fourth Quarter"))) %>%
  ggplot(aes(x = TimeofYear, y = Avg_DepDelay, group = 1)) +
  geom_line() +
  ggtitle("Average Departure Delay based on Quarters") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Quarter") +
    ylab("Average Departure Delay")

#April is the least delayed due to weather on average at 0.37 mins
Best_Time_Weather<- dbGetQuery(base,
  "SELECT Month as Month, ROUND(AVG(alpha.WeatherDelay),2) as Avg_WeatherDelay
  FROM alpha
  GROUP BY Month
  ORDER by Avg_WeatherDelay"
)

#Create a line chart for the average weather delay based on months of the year
Best_Time_Weather %>%
  ggplot(aes(x = Month, y = Avg_WeatherDelay, group = 1)) +
  geom_line() +
  geom_label(aes(label = Month),
              nudge_x = 0.05,
              nudge_y = 0.05,
              size = 3) +
  ggtitle("Average Weather Delay based on Months") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Month") +
    ylab("Average Weather Delay")

#September is the least delayed due to NAS delays on average at 2.22 mins
Best_NAS<- dbGetQuery(base,
  "SELECT Month as Month, ROUND(AVG(alpha.NASDelay),2) as Avg_NASDelay
  FROM alpha
  GROUP BY Month
  ORDER by Avg_NASDelay"
)

#Create a line chart for the average NAS delay based on months of the year
Best_NAS %>%
  ggplot(aes(x = Month, y = Avg_NASDelay, group = 1)) +
  geom_line() +
  geom_label(aes(label = Month),
              nudge_x = 0.05,
              nudge_y = 0.05,
              size = 3) +
  ggtitle("Average NAS Delay based on Months") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Month") +
    ylab("Average NAS Delay")

#Save ram
rm(Best_TimeGivenADDD)
rm(Best_TimeGivenDD)
rm(Best_Time)
rm(Best_Time_Weather)
rm(Best_NAS)
```

Question 1: 
Finding relationship based on Day of Month: Departure & Arrival
```{r}
#3rd is the least delayed at an average of 32.10 mins for departure and
#33.22 mins for arrival 
#This is given that there is arrival and departure delay
Day_DelaysGivenADDD <- dbGetQuery(base,
  "SELECT DayofMonth as DayofMonth, AVG(DepDelay) as Avg_DepDelay, 
  AVG(ArrDelay) as Avg_ArrDelay, Day as Day
  FROM alpha
  WHERE alpha.ArrDelay >0 AND alpha.DepDelay >0
  GROUP BY DayofMonth
  ORDER by Avg_DepDelay"
)

#3rd has the lowest departure delay at 23.06 mins on average 
#This is given that there is only departure delay
Day_DepDelaysGivenDD <- dbGetQuery(base,
  "SELECT DayofMonth as DayofMonth, AVG(DepDelay) as Avg_DepDelay, Day as Day
  FROM alpha
  WHERE alpha.DepDelay >0
  GROUP BY DayofMonth
  ORDER by Avg_DepDelay"
)

#3rd has the lowest arrival delay at 24.08 mins on average
#This is when considering only arrival delay  
Day_ArrDelaysGivenAD <- dbGetQuery(base,
  "SELECT DayofMonth as DayofMonth, AVG(ArrDelay) as Avg_ArrDelay, Day as Day
  FROM alpha
  WHERE alpha.ArrDelay >0
  GROUP BY DayofMonth
  ORDER by Avg_ArrDelay"
)

#3rd is the least delayed at an average of 6.14 mins for departure and 
#3.97 mins for arrival
#This is when considering the whole population
Day_Delays <- dbGetQuery(base,
  "SELECT DayofMonth as DayofMonth, AVG(DepDelay) as Avg_DepDelay, 
  AVG(ArrDelay) as Avg_ArrDelay, Day as Day
  FROM alpha
  GROUP BY DayofMonth
  ORDER by Avg_DepDelay"
)

#Create a box plot showing the arrival delay based on days of the week
AD1 <- Day_Delays %>%
  mutate(Day = factor(Day,
                      levels = c("Mon", "Tues", "Wed", "Thurs",
                                       "Fri", "Sat", "Sun"))) %>%
  ggplot(aes(x = Day, y = Avg_ArrDelay)) +
  geom_boxplot(fill = "skyblue") +
  stat_summary(fun = mean, geom = "point", colour = "red") +
  stat_summary(fun = mean, geom = "text", colour = "red", vjust = 6.1,
               aes(label = paste("Mean:", round(..y..,1)))) +
  ggtitle("Average Arrival Delay based on Day of Week") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Day of Week") +
    ylab("Average Arrival Delay")

#Create a box plot showing the departure delay based on days of the week
DD1 <- Day_Delays %>%
  mutate(Day = factor(Day,
                      levels = c("Mon", "Tues", "Wed", "Thurs",
                                       "Fri", "Sat", "Sun"))) %>%
  ggplot(aes(x = Day, y = Avg_DepDelay)) +
  geom_boxplot(fill = "skyblue") +
  stat_summary(fun = mean, geom = "point", colour = "red") +
  stat_summary(fun = mean, geom = "text", colour = "red", vjust = 6.8,
               aes(label = paste("Mean:", round(..y.., 1)))) +
  ggtitle("Average Departure Delay based on Day of Week") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Day of Week") +
    ylab("Average Departure Delay")

#Create a box plot showing the arrival delay based on days of the month
AD2 <- Day_Delays %>%
  ggplot(aes(x = DayofMonth, y = Avg_ArrDelay)) +
  geom_line(colour = "blue") +
  geom_point(colour = "black") +
  geom_label(aes(label = round(Avg_ArrDelay, 2)),
            nudge_x = 0.25,
            nudge_y = 0.25,
            size = 3) +
  ggtitle("Average Arrival Delay based on Day of Month") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Day of Month") +
    ylab("Average Arrival Delay")

#Create a box plot showing the departure delay based on days of the month
DD2 <- Day_Delays %>%
  ggplot(aes(x = DayofMonth, y = Avg_DepDelay)) +
  geom_line(colour = "blue") +
  geom_point(colour = "black") +
  geom_label(aes(label = round(Avg_DepDelay, 2)),
            nudge_x = 0.25,
            nudge_y = 0.25,
            size = 3) +
  ggtitle("Average Departure Delay based on Day of Month") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Day of Month") +
    ylab("Average Departure Delay")

#Arrange the charts based on a customised order
ggarrange(AD1, DD1, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

#Arrange the charts based on a customised order
ggarrange(AD2, DD2, 
          labels = c("C", "D"),
          ncol = 2, nrow = 1)


#Save ram
rm(Day_DelaysGivenADDD)
rm(Day_DepDelaysGivenDD)
rm(Day_ArrDelaysGivenAD)
rm(Day_Delays)
```

Question 1: 
Finding relationship based on the Month: Arrival & Departure
```{r}
#April is the least delayed at an average of 31.16 mins for departure and 
#32.57 for arrival. 
#This is given that there is arrival and departure delay
Month_DelaysGivenADDDwithDistance <- dbGetQuery(base,
  "SELECT Mth as Month, AVG(DepDelay) as Avg_DepDelay, 
  AVG(ArrDelay) as Avg_ArrDelay, AVG(distance) as Avg_Distance
  FROM alpha
  WHERE alpha.ArrDelay >0 AND alpha.DepDelay >0
  GROUP BY Month
  ORDER BY Avg_DepDelay"
)

#April is the least delayed at an average of 24.98 mins for departure and 
#23.06 mins for arrival
#This is given that there is only departure delay
Month_DelaysGivenDDwithDistance <- dbGetQuery(base,
  "SELECT Mth as Month, AVG(DepDelay) as Avg_DepDelay, 
  AVG(ArrDelay) as Avg_ArrDelay,  AVG(distance) as Avg_Distance
  FROM alpha
  WHERE alpha.DepDelay >0
  GROUP BY Month
  ORDER BY Avg_DepDelay"
)

#September is the least delayed at an average of 4.45 mins for departure and 
#1.85 mins for arrival.
#Average delays seems to increase alongside average distance
Month_DelayswithDistance <- dbGetQuery(base,
  "SELECT Mth as Month, AVG(DepDelay) as Avg_DepDelay, 
  AVG(ArrDelay) as Avg_ArrDelay, AVG(distance) as Avg_Distance
  FROM alpha
  GROUP BY Month
  ORDER BY Avg_DepDelay"
)

#Create a line chart for average departure delays based on the different months
MDD <- Month_DelayswithDistance %>%
  mutate(Month = factor(Month,
                      levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
  ggplot(aes(x = Month, y = Avg_DepDelay, group = 1)) +
  stat_summary(fun = mean, geom = "line", colour = "red") +
  stat_summary(fun = mean, geom = "point", colour = "black") +
  stat_summary(fun = mean, geom = "text", colour = "blue", vjust = 1, hjust = 0.4, size = 3,
               aes(label = paste("Mean:", round(..y.., 1)))) +
  ggtitle("Average Departure Delay based on Month") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Month") +
    ylab("Average Departure Delay")

#Create a line chart for average arrival delays based on the different months
MAD <- Month_DelayswithDistance %>%
  mutate(Month = factor(Month,
                      levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
  ggplot(aes(x = Month, y = Avg_ArrDelay, group = 1)) +
  stat_summary(fun = mean, geom = "line", colour = "red") +
  stat_summary(fun = mean, geom = "point", colour = "black") +
  stat_summary(fun = mean, geom = "text", colour = "blue", vjust = 1, size = 3,
               aes(label = paste("Mean:", round(..y.., 1)))) +
  ggtitle("Average Arrival Delay based on Month") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Month") +
    ylab("Average Arrival Delay")

#Create a line chart for average distance traveled based on the different months
MDist <- Month_DelayswithDistance %>%
  mutate(Month = factor(Month,
                      levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
  ggplot(aes(x = Month, y = Avg_Distance, group = 1)) +
  stat_summary(fun = mean, geom = "line", colour = "red") +
  stat_summary(fun = mean, geom = "point", colour = "black") +
  stat_summary(fun = mean, geom = "text", colour = "blue", vjust = 1, size = 3,
               aes(label = paste("Mean:", round(..y.., 1)))) +
  ggtitle("Average Distance based on Month") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Month") +
    ylab("Average Distance")

#Arrange the charts based on a customised order
ggarrange(MDD, MAD, MDist, 
          labels = c("E", "F", "G"), nrow = 3)

#Save ram
rm(Month_DelaysGivenADDDwithDistance)
rm(Month_DelaysGivenDDwithDistance)
rm(Month_DelayswithDistance)
```

Question 2: 
Do older planes suffer more delays?
```{r}
#Create new column referencing issue date
planes$issue_year = stri_sub(planes$issue_date,-4)

#Join alpha and planes data frame into total_flights
total_flights <- left_join(alpha, planes, by = c("TailNum" = "tailnum"))

#Get age of planes through the subtraction and conversion of issue_year to an integer
total_flights$Age <- total_flights$Year - as.integer(total_flights$issue_year)

#Create a 100% stacked bar chart showing the punctuality of planes based on their age
PlanesAge <- total_flights %>%
  na.omit() %>%
  ggplot(aes(x = Age, fill = Punctuality, labels = Age)) +
    geom_bar(position = "fill") +
    ggtitle("Percentage of Flights that are Puncutal based on Age") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme_classic() +
    xlab("Planes Age") +
    ylab("Percentage of Flights") +
    guides(fill = guide_legend("Punctuality")) 

#Create a 100% stacked bar chart showing the punctuality of planes based on their Manufacturer
Manufacturer <- total_flights %>%
  na.omit() %>%
  ggplot(aes(x = manufacturer, fill = Punctuality)) +
  geom_bar(position = "fill") +
  ggtitle("Percentage of Flights that are Puncutal based on Manufacturer") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Manufacturer") +
  ylab("Percentage of Flights") +
  guides(fill = guide_legend("Punctuality")) 

#Arrange the charts based on a customised order
ggarrange(PlanesAge, Manufacturer, 
          labels = c("H", "I"), ncol = 2)

```

Question 3: 
How does the number of people flying between different locations change over time?
```{r}
#Create a new data frame with specific columns referencing year 2004
Avg_no_of_flights_2004 <- dbGetQuery(base,
  "SELECT Dest as Destination, Year as Year1, COUNT(DepTime) as FrequencyX, 
  SUM(Distance) as Total_DistanceX
  FROM alpha
  WHERE alpha.Year = 2004
  GROUP BY Destination
  Order BY FrequencyX"
)
#Create a new data frame with specific columns referencing year 2005
Avg_no_of_flights_2005 <- dbGetQuery(base,
  "SELECT Dest as Destination, Year as Year2, COUNT(DepTime) as FrequencyY, 
  SUM(Distance) as Total_DistanceY
  FROM alpha
  WHERE alpha.Year = 2005 
  GROUP BY Destination
  Order BY FrequencyY"
)

#Combine the two data frames
Avg_no_of_flights <- left_join(Avg_no_of_flights_2004, Avg_no_of_flights_2005, by = "Destination")

#Replacement to na.omit as it does not work 
Avg_no_of_flights <- Avg_no_of_flights[complete.cases(Avg_no_of_flights),]

#Keep any values greater than 30 under FrequencyX column
#The removal of values under 30 applied for FreqY as well
Avg_no_of_flights <- subset(Avg_no_of_flights, FrequencyX >30,)

#Use arithmetic formulas to get the difference in frequency between the two years
Avg_no_of_flights$Difference_F <- Avg_no_of_flights$FrequencyY - Avg_no_of_flights$FrequencyX

#Use arithmetic formulas to get the percentage change between the two years
Avg_no_of_flights$Percentage_ChangeF <- 
  round(Avg_no_of_flights$Difference_F / Avg_no_of_flights$FrequencyX * 100, 2)

#The change of average number of flights during 2004 and 2005 is 2.72%
print(paste0("The change of average number of flights during 2004 and 2005 is ", round(mean(Avg_no_of_flights[["Percentage_ChangeF"]]),2),"%"))

#Use arithmetic formulas to get the average distance travelled in their respective years
Avg_no_of_flights$Avg_DistX <- 
  (Avg_no_of_flights$Total_DistanceX / Avg_no_of_flights$FrequencyX)
Avg_no_of_flights$Avg_DistY <-
  (Avg_no_of_flights$Total_DistanceY / Avg_no_of_flights$FrequencyY)

#Use arithmetic formulas to get the difference in distance between the two years
Avg_no_of_flights$Difference_D <-round(Avg_no_of_flights$Avg_DistY - Avg_no_of_flights$Avg_DistX,4)

#Use arithmetic formulas to get the percentage change between the two years
Avg_no_of_flights$Percentage_ChangeD <-
  round(Avg_no_of_flights$Difference_D / Avg_no_of_flights$Avg_DistX *100, 2)

#The change of average distance traveled by flights during 2004 and 2005 is 2.37%
print(paste0("The change of average distance of traveled flights during 2004 and 2005 is ", round(mean(Avg_no_of_flights[["Percentage_ChangeD"]]),2),"%"))

#Save ram
rm(Avg_no_of_flights_2004)
rm(Avg_no_of_flights_2005)
rm(Avg_no_of_flights)
```

Question 4: 
Can you detect cascading failures as delays in one airport create delays in others?
Using linear regression to answer how Departure and Arrival delay is correlated.
```{r}
Airport_Delay <- dbGetQuery(base,
  "SELECT Dest as Destination, AVG(DepDelay) as Avg_Departure_Delay, 
  AVG(ArrDelay) as Avg_Arrival_Delay
  FROM alpha
  GROUP BY Destination
  ORDER BY Avg_Departure_Delay
")

Airport_Delay <- na.omit(Airport_Delay)

round(cor(Airport_Delay$Avg_Departure_Delay, Airport_Delay$Avg_Arrival_Delay),3)
#Get the correlation for both departure and arrival delay
#0.824 shows that both data are highly correlated 

dep.arr.lm <- lm(Avg_Arrival_Delay ~ Avg_Departure_Delay, data = Airport_Delay)
#Obtain a linear regression analysis

summary(dep.arr.lm)
#This shows that there is a significant positive relationship between arrival 
#and departure delay, with a 0.7861 (+/- 0.0001) unit increase in departure
#delay for every unit increase in arrival delay

plot(dep.arr.lm)
#The plots shows that departure and arrival delay is so highly correlated that
#the mean line deviates so much that the homoscedasticity assumption is invalid
#https://www.scribbr.com/statistics/linear-regression-in-r/
```

Question 5:
Use the available variables to construct a model that predicts delays.
Using gradient boosting and SVM to predict cascading delays in question 2
```{r}
#Slicing the data set to 10000 random rows
total_flights = total_flights[sample(nrow(total_flights), 10000), ]

#Convert the different variables to columns
total_flights$Punctuality <- factor(total_flights$Punctuality)
total_flights$TimeofYear <- factor(total_flights$TimeofYear)

#Get the roles
num <- nrow(total_flights)

#Get a random sample of 60%
train_set <- sample(num, round(0.6 * num))

#Get the other 40% that is not included in the training set
test_set <- setdiff(1:num, train_set)

#Set the classification task, name it as "total_flights", from "total_flights",
#where the target variable is TimeofYear
NewTask <- TaskClassif$new('total_flights', backend = total_flights, target = "TimeofYear")

#Select 4 variables
NewTask$select(c('Age', 'DepDelay', 'ArrDelay', 'Punctuality'))

#Use Mean Squared Residual to measure the error
measure <- msr('classif.ce')

#Change the factors to numerical
fencoder <- po("encode", method = "treatment",
  affect_columns = selector_type("factor"))

#With tuner, we use grid search 
tuner <- tnr('grid_search')

#Set the number of evaluation we do to 2
terminator <- trm('evals', n_evals = 2)

#Perform gradient boosting - ML
learner_gb <- lrn("classif.xgboost")
gc_gb <- po('imputemean') %>>%
  fencoder %>>% po(learner_gb)
glrn_gb <- GraphLearner$new(gc_gb)
glrn_gb$train(NewTask, row_ids = train_set)
glrn_gb$predict(NewTask, row_ids = test_set)$score() 

#Perform support vector machine - ML
learner_svm <- lrn("classif.svm")
gc_svm <- po('imputemean') %>>% fencoder %>>% po(learner_svm)
glrn_svm <- GraphLearner$new(gc_svm)
glrn_svm$train(NewTask, row_ids = train_set)
glrn_svm$predict(NewTask, row_ids = test_set)$score()

#For reproducible results
set.seed(1) 
lrn_list <- list(
  gc_gb,
  gc_svm
)

#Drawing of the plots  
bm_design <- benchmark_grid(task = NewTask, resamplings = rsmp('cv', folds = 2), learners = lrn_list)
bmr <- benchmark(bm_design, store_models = TRUE)
autoplot(bmr) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + theme_classic()
bmr$aggregate(measure)
```
